---
title: "HW1 613"
author: "Mike Sun"
date: "1/16/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(ineq)
```

```{r, include=FALSE}
# household
dathh2004 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2004.csv")
dathh2005 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2005.csv")
dathh2006 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2006.csv")
dathh2007 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2007.csv")
dathh2008 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2008.csv")
dathh2009 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2009.csv")
dathh2010 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2010.csv")
dathh2011 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2011.csv")
dathh2012 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2012.csv")
dathh2013 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2013.csv")
dathh2014 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2014.csv")
dathh2015 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2015.csv")
dathh2016 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2016.csv")
dathh2017 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2017.csv")
dathh2018 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2018.csv")
dathh2019 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/dathh2019.csv")
```

```{r,include=FALSE}
#individual
datind2004 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2004.csv")
datind2005 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2005.csv")
datind2006 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2006.csv")
datind2007 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2007.csv")
datind2008 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2008.csv")
datind2009 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2009.csv")
datind2010 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2010.csv")
datind2011 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2011.csv")
datind2012 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2012.csv")
datind2013 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2013.csv")
datind2014 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2014.csv")
datind2015 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2015.csv")
datind2016 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2016.csv")
datind2017 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2017.csv")
datind2018 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2018.csv")
datind2019 <- read.csv("~/Desktop/Fourth_Semester/613/HW1/Data/datind2019.csv")
```


$$ Excercise 1 $$

Number of households surveyed in 2007: 

```{r}
length(dathh2007$idmen)
```

Number of households with marital status “Couple with kids” in 2005:

```{r}
nrow(filter(dathh2005, mstatus == "Couple, No kids"))
```

Number of individuals surveyed in 2008:

```{r}
length(datind2008$idind)
```

Number of individuals aged between 25 and 35 in 2016:

```{r}
length(filter(datind2016, age >= 25 & age <=35))
```

Cross-table gender/profession in 2009:

```{r}
gender_prof_table_09 = table(datind2009$profession, datind2009$gender)
gender_prof_table_09
```


Distribution of wages in 2005 and 2019. Report the mean, the standard deviation, the inter-decile ratio D9/D1 and the Gini coefficient:

```{r}
datind2005_wageclean = datind2005 %>% filter(!is.na(wage))
datind2019_wageclean = datind2019 %>% filter(!is.na(wage))
ggplot(data=datind2005_wageclean, aes(x=wage))+geom_histogram(binwidth=10000)
ggplot(data=datind2019_wageclean, aes(x=wage))+geom_histogram(binwidth=10000)
```

```{r}
mean(datind2005_wageclean$wage)
mean(datind2019_wageclean$wage)

sd(datind2005_wageclean$wage)
sd(datind2019_wageclean$wage)

D9_1_2005 = quantile(datind2005_wageclean$wage, 0.9)/quantile(datind2005_wageclean$wage, 0.1)
D9_1_2019 = quantile(datind2019_wageclean$wage, 0.9)/quantile(datind2019_wageclean$wage, 0.1)

Gini(datind2005_wageclean$wage)
Gini(datind2019_wageclean$wage)
```

Distribution of age in 2010. Plot an histogram. Is there any difference between men and women?

```{r}
datind2010_ageclean = datind2010 %>% filter(!is.na(age))

ggplot(datind2010_ageclean) + aes(age) + geom_histogram(binwidth=5) + 
  ggtitle("2010 Population Count by Age with Bin size = 4")

male_age_2010 <- datind2010_ageclean %>% filter( gender == "Male" ) %>% ggplot( aes(x=age)) +
    geom_histogram( binwidth=4, fill="#69b3a2", color="#69b3a2") + 
  ggtitle("2010 Male Count by Age with Bin size = 4")

female_age_2010 <- datind2010_ageclean %>% filter( gender == "Female" ) %>% ggplot( aes(x=age)) +
    geom_histogram( binwidth=4, fill = "blue", color="blue") + 
  ggtitle("2010 Female Count by Age with Bin size = 4")

male_age_2010
female_age_2010

```

No noticable difference based on the graphs.


Number of individuals in Paris in 2011:

```{r}
Pop_ind = left_join(dathh2011,datind2011,by="idmen")
Paris_ind = Pop_ind %>% filter(location == "Paris")
nrow(Paris_ind)
```

\newpage

$$ Excercise 2 $$

Read all individual datasets from 2004 to 2019. Append all these datasets:

```{r}
datind_full = unique(rbind(datind2004,datind2005,datind2006,datind2007,datind2008,datind2009,
                           datind2010,datind2011,datind2012,datind2013,datind2014,datind2015,
                           datind2016,datind2017,datind2018,datind2019))
```

Read all household datasets from 2004 to 2019. Append all these datasets:

```{r}
dathh_full = unique(rbind(dathh2004,dathh2005,dathh2006,dathh2007,dathh2008,
                          dathh2009,dathh2010,dathh2011,dathh2012,dathh2013,
                          dathh2014,dathh2015,dathh2016,dathh2017,dathh2018,dathh2019))
```

List the variables that are simultaneously present in the individual and household datasets:

```{r}
col_check = colnames(dathh_full) %in% colnames(datind_full)

i=1
while (i <= length(col_check)){
  if (col_check[i] == "TRUE"){
    print(colnames(dathh_full)[i])
  }
  i= i+1
}

```

Number of households in which there are more than four family members:

```{r}
Full_Data = unique(left_join(dathh_full,datind_full,by=c("idmen","year")))
```

```{r}
idmen_count <- Full_Data %>% group_by(idmen,year) %>% mutate(count_4 = n()) %>% filter(count_4 > 4)
```

```{r}
length(idmen_count$idmen)
```



Number of households in which at least one member is unemployed:

```{r}
Unemployed = Full_Data %>% group_by(idmen, year, empstat) %>% filter(empstat == "Unemployed")

length(unique(Unemployed$idmen))
```

Number of households in which at least two members are of the same profession:

```{r}
Same_Profession = Full_Data %>% group_by(idmen, year, profession) %>% 
  mutate(n_member = n()) %>% filter(n_member > 2)

length(unique(Same_Profession$idmen))
```

Number of individuals in the panel that are from household-Couple with kids:

```{r}
Couple_kids = Full_Data %>% group_by(idmen,year,mstatus) %>% filter(mstatus == "Couple, with Kids")

length(unique(Couple_kids$idind))
```

Number of individuals in the panel that are from Paris:

```{r}
Paris = Full_Data %>% group_by(idind,year,location) %>% filter(location == "Paris")

length(unique(Paris$idind))
```

Find the household with the most number of family members. Report its idmen:

```{r}
max_idmen = idmen_count %>% filter(count_4 == max(idmen_count$count_4))
unique(max_idmen$idmen)
```

Number of households present in 2010 and 2011:

```{r}
dathh_1011 = intersect(dathh2010$idmen,dathh2011$idmen)
length(dathh_1011)
```

\newpage

$$ Exercise 3 $$

Find out the year each household enters and exit the panel. Report the distribution of the time spent in the survey for each household.

```{r}
entry_exit <- Full_Data %>% group_by(idmen) %>% summarize(entry = min(year), exit = max(year)) %>%
  mutate(time_spent = (exit - entry +1))
```

```{r}
ggplot(entry_exit, aes(x=time_spent)) + geom_histogram(bins = 30) + ggtitle("Time Spent Histogram")
```



Based on datent, identify whether or not a household moved into its current dwelling at the year of survey. Report the first 10 rows of your result and plot the share of individuals in that situation across years:

```{r}
hh_year = Full_Data %>% mutate(move_in_check = ifelse(datent == year,1,0))
```

```{r}
head(hh_year, 10)
```

```{r}
hh_year_ratio = hh_year %>% group_by(year) %>% 
  summarise(total_c = n(), move_c = sum(move_in_check,na.rm = T), m_ratio = move_c/total_c)
```

```{r}
ggplot(hh_year_ratio, aes(y=m_ratio,x=year)) + geom_line() + 
  ggtitle("Move in by Year") + xlab("Year") + ylab("Share of Moved in Popualtion")
```


Based on myear and move, identify whether or not household migrated at the year of survey. Report the first 10 rows of your result and plot the share of individuals in that situation across years:

```{r}
summary(Full_Data)
```


```{r}
migrate_year = Full_Data %>% 
  mutate(mi_check = ifelse(is.na(move) == F & move ==2, 1, ifelse(year == myear, 1 , 0)))

head(migrate_year, 10)
```

```{r}
migrate_year_share = migrate_year %>% group_by(year) %>% 
  summarise(year_count = n(), mi_count = sum(mi_check,na.rm = TRUE)) %>% 
  mutate(mi_share = mi_count/year_count)
```

```{r}
ggplot(migrate_year_share, aes(y=mi_share,x=year)) + geom_line() + 
  ggtitle("Migrate by Year") + xlab("Year") + ylab("Share of Migrated Population")
```


Mix the two plots you created above in one graph, clearly label the graph. Do you prefer one method over the other? Justify:

```{r, include=FALSE}
by_plot = left_join(migrate_year_share, hh_year_ratio, by = "year") %>% 
  select(year, mi_share, m_ratio)
```

```{r}
ggplot(by_plot, aes(x=year)) + 
   geom_line( aes(y=m_ratio),color="red") +  geom_line( aes(y=mi_share), color="blue") +
  
  scale_y_continuous(name = "Average Time Spent",sec.axis = sec_axis(trans=~.*1,,name="Migration Share")) +
  
  theme(
    axis.title.y = element_text(color = "red", size=13),
    axis.title.y.right = element_text(color = "blue", size=13)
  )

```

I prefer the time spent graph since it is straightfoward to understand. Moreover, migration share data has many missing values. In fact, after 2014, we do not have data on myear, and this inconsistency in measurement partially contributes to the jump after 2015. 


For households who migrate, find out how many households had at least one family member changed his/her profession or employment status:

```{r}
prof_emp_c = migrate_year %>% filter(mi_check == 1) %>% 
  group_by(idmen,profession,empstat) %>% mutate(check = n()) %>% 
  mutate(ifelse(check != 1, 1,0)) %>% filter(check == 1)
```

```{r}
length(prof_emp_c$idmen)
```

\newpage

$$Exercise 4$$

Compute the attrition across each year, where attrition is defined as the reduction in the number of indi- viduals staying in the data panel. Report your final result as a table in proportions.


```{r}
years <- 2004:2018
ratio <- seq(2005, 2019, by = 1)
attrition <- seq(2005, 2019, by = 1)
ind_c <- seq(2005, 2019, by = 1)
n = 1
for (i in years) {
  j <- i + 1
  attrition[n] <- length(setdiff(Full_Data[Full_Data$year == i,]$idind, 
                              Full_Data[Full_Data$year == j,]$idind))
  ind_c[n] <- length(Full_Data[Full_Data$year == i,]$idind)
  ratio[n] <- attrition[n]/ind_c[n]
  n = n + 1
}

attrition_ratio <- data.frame(year = 2005:2019, attrition_ratio = ratio)

attrition_ratio
```







